DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-01-31 paper/intro.tex
Minor changes to abstract and intro

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is, in a sense, an ideal real-time
workload as the amount of computation needed to compete a single
inference is fixed at the DNN architecture design time and does not
change at runtime over different input (e.g., different image frame).
This means that it is feasible to achieve highly predictable and
tightly bounded worst-case execution time estimation.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's DNN-based
control loop is feasible on today's embedded computing platforms, even 
as inexpensive as Raspberry Pi 3. We find that the control loop completes 
under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two cores of
the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; we can achieve 20 Hz control frequency using
just one core (out of four) of the Raspberry Pi 3's quad-core CPU.
Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance inference as we observe up to 11.6X control loop
execution time increase, when memory performance intensive
applications are co-scheduled on idle CPU cores.
We find, however, that DNN inferencing is not sensitive to cache
space; we show that applying cache partitioning has negligible impact
in providing timing isolation on consolidated scenarios. This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee its memory performance
are much more important.

>>>>

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is, in a sense, an ideal real-time
workload as the amount of computation needed to compete a single
inference is fixed at the DNN architecture design time and does not
change at runtime over different inputs (e.g., different image frame).
This means that it is feasible to achieve highly predictable and
tightly bounded worst-case execution time estimation.

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's DNN-based
control loop is feasible on today's embedded computing platforms, even 
as inexpensive as the Raspberry Pi 3. We find that the control loop 
completes in under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two 
cores of the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; we can achieve 20 Hz control frequency using
just one core of the Raspberry Pi 3's quad-core CPU.
Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance inference as we observe up to an 11.6X control loop
execution time increase, when memory performance intensive
applications are co-scheduled on idle CPU cores.
We find, however, that DNN inferencing is not sensitive to cache
space; we show that applying cache partitioning has negligible impact
in providing timing isolation on consolidated scenarios. This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee its memory performance
are much more important.

### 2018-01-31 paper/main.tex
Minor changes to abstract and intro

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{Michael G. Bechtel}
\author[2]{Elise McEllhiney}
\author[3]{Minje Kim}
\author[4]{Heechul Yun}
\affil[1]{University of Kansas, Lawrence, United States\\
  \texttt{mbechtel@ku.edu}}
\affil[2]{University of Kansas, Lawrence, United States\\
  \texttt{e908m429@ku.edu}}
\affil[3]{Indiana University, Bloomington, United States\\
  \texttt{e908m429@ku.edu}}
\affil[4]{University of Kansas, Lawrence, United States\\
  \texttt{heechul.yun@ku.edu}}
\authorrunning{M.\,G. Bechtel, E. McEllhiney, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

@@

\maketitle
\thispagestyle{empty}
\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses the exact same network
architecture---8 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

>>>>

%% Please provide for each author the \author and \affil macro, even when authors have the same affiliation, i.e. for each author there needs to be the  \author and \affil macros
\author[1]{Michael G. Bechtel}
\author[2]{Elise McEllhiney}
\author[3]{Minje Kim}
\author[4]{Heechul Yun}
\affil[1]{University of Kansas, Lawrence, United States\\
  \texttt{mbechtel@ku.edu}}
\affil[2]{University of Kansas, Lawrence, United States\\
  \texttt{e908m429@ku.edu}}
\affil[3]{Indiana University, Bloomington, United States\\
  \texttt{minje@indiana.edu}}
\affil[4]{University of Kansas, Lawrence, United States\\
  \texttt{heechul.yun@ku.edu}}
\authorrunning{M.\,G. Bechtel, E. McEllhiney, H. Yun} %mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et. al.'

@@

\maketitle
\thispagestyle{empty}
\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses almost the exact same 
network architecture---8 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}
